---
title: "TMDB Box revenue prediction"
author: "Nguyen Bao Long"
date: "8/19/2019"
output:
  pdf_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
---

\newpage

```{r setup, include=FALSE}
# set up environment, standardize figure output
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = TRUE, fig.height = 3.5, fig.width = 5,
                      fig.align = "center")
```

```{r, include=FALSE}
if(!require(matrixStats)) install.packages("matrixStats", 
                   repos = "http://cran.us.r-project.org", 
                   dependencies = TRUE)
if(!require(Amelia)) install.packages("Amelia", 
                   repos = "http://cran.us.r-project.org", 
                   dependencies = TRUE)
if(!require(gridExtra)) install.packages("gridExtra", 
                   repos = "http://cran.us.r-project.org", 
                   dependencies = TRUE)
if(!require(tidyverse)) install.packages("tidyverse", 
                   repos = "http://cran.us.r-project.org", 
                   dependencies = TRUE)
if(!require(caret)) install.packages("caret", 
                   repos = "http://cran.us.r-project.org", 
                   dependencies = TRUE)
if(!require(kableExtra)) install.packages("kableExtra", 
                   repos = "http://cran.us.r-project.org", 
                   dependencies = TRUE)
if(!require(TMDb)) install.packages("TMDb", 
                   repos = "http://cran.us.r-project.org", 
                   dependencies = TRUE)
if(!require(GGally)) install.packages("GGally", 
                   repos = "http://cran.us.r-project.org", 
                   dependencies = TRUE)
```


```{r import library, include=FALSE}
library(tidyverse)
library(caret)
library(matrixStats)
library(grDevices)
library(gridExtra)
library(Amelia)
library(DataExplorer)
library(knitr)
library(kableExtra)
library(lubridate)
library(corrplot)
library(TMDb)
library(GGally)
```


\newpage

**Abstract** This document belongs to the second project - Choose Your Own of the course HarvardX: PH125.9x Data Science: Capstone project. Inspired by the difficulty and challenging in Kaggle's competitions, and my expectation to practice as much as possible data science techniques I have learned in the HarvardX - Data Science program, in this project I used the dataset from Kaggle TMDb Box Office Competition, where competitors were presented with metadata on over 7,000 past films from The Movie Database to try and predict their overall worldwide box office revenue. Furthermore, this competition allowed competitors to use additional publicly dataset for modeling, therefore it's good opportunity to practice the web scrapping techniques. This document describe my solution include data preparation, data wrangling, data visualization, model development and its performance evaluation. The main R packages used in this project were tidyverse for data wrangling, knitr & kableExtra and ggplot2 & gridExtra for data visualization, revest, TMDb for web scraping, caret for modeling and modeling performance evaluation.

# **Introduction**

## **Background**

### **TMDB Box Office Competition**

*"Can you predict a movie's worldwide box office revenue?"*

On February $9^{th}, 2019$ Kaggle launch a competition, using the TMDB Box Office dataset to predict a movie's worldwide box office revenue. This competition was end by May $31^{th}$, 2019 which total 1398 teams, 1618 competitors and 19,034 entries. Here come the introduction from Kaggle:

*"In a world… where movies made an estimated $41.7 billion in 2018, the film industry is more popular than ever. But what movies make the most money at the box office? How much does a director matter? Or the budget? For some movies, it's "You had me at 'Hello.'" For others, the trailer falls short of expectations and you think "What we have here is a failure to communicate."*

*In this competition, you're presented with metadata on over 7,000 past films from The Movie Database to try and predict their overall worldwide box office revenue. Data points provided include cast, crew, plot keywords, budget, posters, release dates, languages, production companies, and countries. You can collect other publicly available data to use in your model predictions, but in the spirit of this competition, use only data that would have been available before a movie's release."*

**Competition evaluation**

To evaluate results in Kaggle's competition, competitors must develop machine learning model to predict the international box office revenue for each movie. For each id in the test set, competitors must predict the value of the revenue variable. 

Submissions are evaluated on Root-Mean-Squared-Logarithmic-Error (RMSLE) between the predicted value and the actual revenue. Logs are taken to not overweight blockbuster revenue movies.

Since this competition completed on May $31^{st}$, following is the final public leaderboard.

```{r, fig.width=8, include=FALSE}
# import leaderboard data
url_leaderboard <- url("https://raw.githubusercontent.com/billynguyenlss/TMDB-Box/master/data/tmdb-publicleaderboard.csv")
tmdb_leaderboard <- read.csv(url_leaderboard, na.strings=c("", '#N/A', '[]', '0'))
tmdb_leaderboard$SubmissionDate <- ymd_hms(tmdb_leaderboard$SubmissionDate)
```

```{r, include=FALSE}
final_board <- tmdb_leaderboard %>% group_by(TeamName) %>%
  summarize(submitted_times = n(),
            best_score = min(Score)) %>% arrange(best_score)
```

```{r, include=FALSE}
top_10 <- final_board[140,3]
```

```{r, fig.width=8}
final_board %>% 
  ggplot(aes(best_score)) +
  geom_histogram(fill = "steel blue", binwidth = 0.1) +
  scale_x_continuous(minor_breaks = seq(0, 30, 1)) +
  geom_vline(xintercept = 1.7068, color = "red") +
  geom_text(aes(x = 5, y = 240, label = "top 10% (RMSLE < 1.7068)"), color = "red") +
  labs(title = "TMDB Box office competition final leaderboard",
       caption = "Data source from Kaggle's TMDB Box office competition public leaderboard")
```

**Noted:** another point that unfortunetely, even Kaggle's competition rules mentioned *"use only data that would have been available before a movie's release"*, but I found that many competitors used the vote count and vote average in their model, which should be available after the movie's release. Therefore in this project, I did not submit the results to the competition's leaderboard for a benchmarking.

## **Data overview**

### **Competition's dataset**

The dataset used in this competition has been collected from the Movies Database (https://www.themoviedb.org/). The movie details, credits and keywords have been collected from the TMDB Open API. This competition uses the TMDB API but is not endorsed or certified by TMDB. Their API also provides access to data on many additional movies, actors and actresses, crew members, and TV shows.

This data contain 7398 movies and a variety of metadata. Movies are labeled with id. Data points include cast, crew, plot keywords, budget, posters, release dates, languages, production companies, and countries. The dataset was subset to 2 dataset, the `train.csv` has 3000 movies and `test.csv` with 4398 movies.

### **Additional dataset**

**The Movies Database API**

Since this competition allowed for publicly available data, I retrieved the additional data from TMDB website (https://www.themoviedb.org/?language=en-US).

The instruction to retrieve data from TMDb website is reference from following link (http://www.planetanalytics.in/2017/05/how-to-extract-movie-data-from-movie.html).

The TMDb packages was also available in R libraries (https://cran.r-project.org/web/packages/TMDb/index.html) which you can retrieve other features for this competition. 

**Wikipedia**

The second additional data was scrapped from wikipedia, using `rvest` package. The wikipedia data was used to replace the NA values in the original Kaggle's dataset, which would be explained more detailed in post-sections.

## **My objective**

My objective are:

1) To develop a machine learning model to predict the internaltion box office revenue and reduce RMSLE as much as possible.

2) To practice data exploratory analysis techniques which I have learned in this course:
  * Text mining using `stringr`, 
  * Data visualization using `ggplot2`, `gridExtra`, `knitr`, `kableExtra`,
  * Summarize and transform data using `dplyr`, 
  * Web scrapping using `rvest`,
  * Machine learning modeling using `caret`.

3) To improve my business writing and data science reporting skills, using R and `rmarkdown`.

## **Project methodology**

This project has 4 main steps:

+ Step 1 **data exploratory analysis**: exploring, cleaning, visualizing the data to have an overview with-in and between the variables, what's insights gained after analysis.

+ Step 2:**features engineering**  This section describe the new potential features adding to the machine learning model to improve performance. Inputs for this step are insights gained in previous step, and some best score kernels in Kaggle. Below are some kernel which inspired me to complete my report.

  * Zero92's kernel 
  
  (RMSLE = 1.6998) https://www.kaggle.com/zero92/tmdb-prediction
  
  * tavoosi's kernel 
  
  (RMSLE = 1.95841) https://www.kaggle.com/tavoosi/predicting-box-office-revenue-with-random-forest

+ Step 3: **modeling and performance evaluation**. This section describe the modeling development process, comparing different machine learning method and selecting the best method for final modeling. The metric to evaluate model performance is RMLSE which is same standard with the Kaggle Movies Box Office competition. The RMLSE are calculated by following formula:

\begin{center}

$RMLSE = \sqrt{\frac{\sum_{i = 1}^{n} (log y_{i} - log \hat{y}_{i})^{2}}{n}}$

\end{center}

With:
 * $y_{i}$ the true revenue of movie i
 * $\hat{y}_{i}$ the predicted revenue of movie i

```{r}
# write function RMLSe to evaluate modeling performance
RMLSE <- function(predicted_revenue, true_revenue){
  sqrt(mean((log(true_revenue)- log(predicted_revenue))^2))
}
```

* Step 4: **validation on the test set**. This section describe the peformance of the final model from previous step, using the test dataset. The true revenue for this step is received from the additional TMDB data. The metric to evaluate final performance was RMLSE as well.

# **Data exploratory analysis**

## **Data preparation**

All datas which were used in this project, are able to download from my github repo.

**Kaggle TMDB Box Office Competition's dataset**

```{r}
# download raw data from my github repo
url_train <- 
  url("https://raw.githubusercontent.com/billynguyenlss/TMDB-Box/master/data/train.csv")
train_set <- read.csv(url_train, na.strings=c("", '#N/A', '[]', '0'))

url_test <- 
  url("https://raw.githubusercontent.com/billynguyenlss/TMDB-Box/master/data/test.csv")
test_set <- read.csv(url_test, na.strings=c("", '#N/A', '[]', '0'))

url_sample_submission <- 
  url("https://raw.githubusercontent.com/billynguyenlss/TMDB-Box/master/data/sample_submission.csv")
sample_submission <- read.csv(url_sample_submission, na.strings=c("", '#N/A', '[]', '0'))
```

Combine train and test dataset to one data to save time for data preprocessing and transformation.

```{r combine train test data, warning=FALSE}
# combine train_set and test_set to reduce time to pre-processing data
train_set <- train_set %>% mutate_if(is.factor, as.character)
test_set <- test_set %>% mutate_if(is.factor, as.character)
df <- bind_rows(train_set, test_set)
```

**TMDb's additional dataset**

Due to very long processing time to download data from TMDb website, I did not include the code in this report, but you can follow the instruction from previous sections or download data direct from my github repo.

```{r}
# download additional data from my github repo
url_additional <- url("https://raw.githubusercontent.com/billynguyenlss/TMDB-Box/master/data/personal%20additional%20data/full_additional_features_2.csv")
addition_data <- read.csv(url_additional, na.strings=c("", '#N/A', '[]', '0'))
```

Because the true revenue was available on the TMDb dataset, I extract it to a vector named `test_y` for final modeling performance evaluation.

```{r, warning=FALSE}
# extract the true_rating value from additional data
test_y <- test_set %>% left_join(addition_data, by = "imdb_id") %>%
  mutate(new_revenue = ifelse(is.na(new_revenue), median(new_revenue, na.rm = T), new_revenue)) %>%
  pull(new_revenue)
```

**Wikipedia's dataset**

Describe in later section.

## **Data overview**

The **train_set** have 3000 observations of 23 variables.

```{r}
dim(train_set)
```

The **test_set** has 4398 observations of 22 variables.

```{r}
dim(test_set)
```

Lets take a glimpse at Kaggle's dataset to get a feel of how it looks like.

```{r}
glimpse(df)
```

We could classify and summarize the variables to data type continuous, discrete and date/time. The first column `id` was ordering number, `imdb_id` was the identified number in IMDB system, so I did't include them in below summarize table. But please be noticed that the `imdb_id` was important to retrieve additional data from TMDb database website.

Continuous variables    Discrete variables      Date/time     
----------------------  ----------------------  ------------- 
budget                  belongs_to_collection   release_date
popularity              genres
runtime                 homepage
revenue                 original_language
                        original_title
                        overview
                        poster_path
                        production_companies
                        production_contries
                        spoken_languages
                        tagline
                        Keywords
                        cast
                        crew
----------------------  ----------------------  ------------- 

**TMDb additional data**

This data set have 7398 observations of 13 variables. The `vote_count` and `vote_average` were available after the movie had been released. I include them in my addition_data to have a feel on how it look like, but not include in my machine learning model.

```{r}
glimpse(addition_data)
```

**NA value**

Missing values must be dropped or replaced in order to draw correct conclusion from the data. Let's take a look on the NA value of each variables.

```{r, fig.height=5}
plot_missing(train_set)
```

`belongs_to_collection` and  `homepage` have high ratio of NA value, more than 60%.
`budget` have 27.07% NA values.
Other `tagline`, `Keywords`, `production_companies` have lower NA value ratio, less than 20%.

# **Data exploratory analysis**

## **Revenue**

Summary table of revenue:

```{r}
average_revenue <- mean(df$revenue, na.rm = TRUE)
summary(train_set$revenue)
```

Histogram and boxplot are great plot to visualize a group of continuous variables.

```{r, fig.width=8, fig.height=3.5}
grid.arrange(
train_set %>% ggplot(aes(revenue)) +
  geom_histogram(fill = "steel blue", color = "white") +
  geom_vline(aes(xintercept = average_revenue), color = "red") +
  labs(title = "frequency of revenue",
       caption = "Figure - histogram using ggplot2 in train_set"),
train_set %>% ggplot(aes(y = revenue)) +
  geom_boxplot(fill = "steel blue") +
  labs(title = "boxplot of revenue",
       caption = "Figure - boxplot using ggplot2 in train_set") +
  coord_flip(),
ncol = 2)
```

Not many movies have high revenue.
`r 100*round(mean(train_set$revenue <= average_revenue),3)`
% movies in the training dataset have revenue lower than the average revenue. Only
`r 100 - 100*round(mean(train_set$revenue <= average_revenue),3)`
% movies have revenue greater than the average revenue.

Let's see performance of the first naive model using the average revenue as the predicted revenue, compared to the true revenue in train dataset.

```{r}
naive_RMLSE <- RMLSE(average_revenue, train_set$revenue)
naive_RMLSE
```

### **release date**

**What's data tell us?**

The current class of release_date was "chr" (character), so we firstly convert it to type "date and time".

```{r}
df$release_date <- mdy(df$release_date)
summary(df$release_date)
```

The time frame was really wide, the soonest released date was 1969-01-01 and the latest is "2068-12-30".

The latest release_date "2068-12-30" was impossible. There were total
`r df %>% filter(release_date >= mdy("8-27-2019")) %>% nrow`
days later than the time I was doing this report on August $27^{th}$, 2019.
We also havd 1 NA value in the `release_date`. 

To correct the `release_date` and replace NA value, I used the release date in TMDb additional data set. Now we merged the addition data to the `df` dataset.

```{r, warning=FALSE}
# left joining the addition data into df
df <- df %>% left_join(addition_data[,c(1,2,3,7:13)], by = "imdb_id")
df$new_release_date <- as.character(df$new_release_date)

# replace wrong release data
df <- df %>% 
  mutate(new_release_date = mdy(new_release_date)) %>%
  mutate(release_date = ifelse(release_date >= ymd("2019-05-31"),
                               new_release_date, release_date)) %>%
  mutate(release_date = ifelse(is.na(release_date),
                               new_release_date, release_date)) %>%
  mutate(release_date = as_date(release_date))
```

```{r}
summary(df$release_date)
```

We could see the updated time frame as following. The soonest release date was "1921-02-06" and the latest release date was "2018-08-01", there were around 97 years. Please be noticed that many events happened in the world during this time, world war II, economic crisis, technology change, movie's audience interest... could impacted direct to the currency, exchange rate and indirect to movies box office industry.

Because we have total
`r length(levels(factor(df$release_date)))`
different release date,
to simplify this variable and visualize the change during time, I calculated the `release_year` and `release_month`.

```{r}
# calculate the release year and release month

df$release_year <- year(df$release_date)
df$release_month <- month(df$release_date)
```

```{r, fig.width=7.5, fig.height=3.5}
# visualize the different number of movies by release year and release month
grid.arrange(
  df %>% 
    ggplot(aes(release_year)) +
    geom_bar(fill = "steel blue") +
    labs(title = "Number of movies by release year"),
  
  df %>% ggplot(aes(release_month)) +
    geom_bar(fill = "steel blue", color = "white") +
    labs(title = "Number of movies by release month"),
  ncol = 2
)
```

The movies box office industry growth slowly before decade 1980s , after this time it's continuous growing every year. We could see
25% movies released in total
`r 1993 - 1921`
years from 1921 to 1993, but the next 25% movies released in only
`r 2001 - 1993`
years from 1993 to 2001.
50% remain movies released in 
`r 2018 - 2001`
after 2001.

Because the number of released movies increased every year, next I explored the change across time of budget and revenue to have a feel on it.

```{r,fig.width=8}
df <- df %>% mutate(revenue_status = ifelse(is.na(revenue),"unknown revenue","known revenue")) %>%
  mutate(revenue_status = factor(revenue_status))
grid.arrange(
    df %>% 
    group_by(release_year, revenue_status) %>%
    summarize(total_budget = sum(budget, na.rm = T)) %>%
    ggplot(aes(release_year, total_budget, fill = revenue_status)) +
    geom_col(color = "white") +
    scale_fill_manual(values = c("grey","steel blue")) +
    theme_classic()+
    theme(legend.position = "none") + ylim(0,1.5e10)+
      labs(title = "budget change by time"),
  
  df %>% group_by(release_year) %>%
    summarize(total_known_revenue = sum(revenue, na.rm = T)) %>%
    ggplot(aes(release_year, total_known_revenue)) +
    geom_col(color = "white", fill = "steel blue") +
    theme_classic()+
    theme(legend.position = "none")+ ylim(0,1.5e10) +
    labs(title = "known revenue change by time"),
  
  ncol = 2)
```

The trend across time of total budget and known revenue from train_set were similar to the total number of movies release every year.  

## **Budget, popularity, runtime**

The remain continuous variables included budget, popularity, runtime.

```{r, fig.height=5, fig.width=8, warning=FALSE}
# overview graph on budget, revenue, runtime, popularity, ROI
grid.arrange(
  train_set %>% ggplot(aes(budget)) +
    geom_histogram(fill = "steel blue", color = "white"),
  
  train_set %>% ggplot(aes(revenue)) +
    geom_histogram(fill = "steel blue", color = "white"),
  
  train_set %>% ggplot(aes(runtime)) +
    geom_histogram(fill = "steel blue", color = "white"),
  
  train_set %>% ggplot(aes(popularity)) +
    geom_histogram(fill = "steel blue", color = "white"),
  
  # histogram ROI (filter out budget less than 50000)
  df %>% filter(budget > 50000) %>%
    mutate(ROI = (revenue - budget)/budget) %>%
    ggplot(aes(ROI)) +
    geom_histogram(fill = "steel blue", color = "white"),
  
  ncol = 3
)
```

```{r}
train_set %>% select(budget, revenue, runtime, popularity) %>%
  summary
```

The minimum budget 
`r min(train_set$budget, na.rm = TRUE)` 
and minimum revenue 
`r min(train_set$revenue, na.rm = TRUE)`
were impossible. Those numbers are outliers and have negative impact on our predicting models. There were also 14 NA values in runtime.

### **Budget**

To explore the budget and revenue, let's create a summary table as following:

```{r}
train_set %>% select(budget, revenue) %>%
  mutate(budget_status = ifelse(is.na(budget), "missing budget",
                                ifelse(budget <=1000,"low budget","normal budget")),
         revenue_status = ifelse(revenue <= 1000,"low revenue","normal revenue")) %>%
  group_by(budget_status, revenue_status) %>%
  summarize(count = n(),
            percent = 100*round(n()/nrow(train_set),3)) %>%
  mutate(remarks = ifelse(count == 5,"wrong budget",
                   ifelse(count == 34, "NA impact",
                   ifelse(count == 778, "NA impact",
                   ifelse(count == 10, "wrong revenue","-"))))) %>%
  kable(caption = "Summarized table of abnormal budget and revenue") %>%
  kable_styling(latex_options = c("HOLD_position"),
                bootstrap_options = c("striped", "hover"),
                full_width = F, position = "float_right")
```

The biggest impact come from NA value budgets. Because this competition allowed to use the additional data, I scrapped data from wikipedia to correct the budget data, instead of using random forest or other machine learning model to predict the missing budget number. 

Because of long processing time to scrapping and pre-processing, I copied but not allowed the scrapping code running in my report. You can refer to the scrapping code in Appendix section or download the completed data from my github repo. The code to download and import the additional data was as following:

```{r}
# download data from my github repo
url_wikipedia_budget <- url("https://raw.githubusercontent.com/billynguyenlss/TMDB-Box/master/wikipedia_us_budget.csv")
wikipedia_budget <- read.csv(url_wikipedia_budget, na.strings=c("", '#N/A', '[]', '0'))

# to assure the final budget as numeric type
wikipedia_budget$final_budget <- as.numeric(wikipedia_budget$final_budget)

# check if any NA value in the additional budget data
mean(is.na(wikipedia_budget$final_budget))
```

Now joining the wikipedia to our data. 

```{r}
# joining addition wikipedia budget
df <- df %>% mutate(imdb_id = as.factor(imdb_id)) %>%
  left_join(wikipedia_budget[,c(1,2,5)],by = "imdb_id")
#replace low budget value
df <- df %>% 
  mutate(budget = ifelse(budget <= 1000 & revenue > 10000, final_budget, budget))
# replace NA budget value by wikipedia budget
df <- df %>% mutate(budget = ifelse(is.na(budget),
                                    final_budget, budget))
mean(is.na(df$budget))
```

The NA value was reduced from 27.07% to 3.12%. This NA values come from movies which the budget currency were not USD($), for example Euro, Crore, PHP, JPY...

Due to high complexity to convert budget from different currencies to usd in a wide time frame (from 1921 to 2018), I replaced the NA values by median budget number per release year.

```{r}
# create summary budget table by release year
budget_by_year <- df %>%
  group_by(release_year) %>%
  summarize(avg_budget = mean(budget, na.rm = T),
            median_budget = median(budget, na.rm = T))

# replace NA budget by median_budget value
df <- df %>% left_join(budget_by_year, by = "release_year") %>%
  mutate(budget = ifelse(is.na(budget), median_budget,
                         budget))

# re-check NA value in budget
mean(is.na(df$budget))
```

### **Runtime**

We replace the NA values in `runtime` by median value.

```{r}
df$runtime[is.na(df$runtime)] <- median(df$runtime, na.rm = T)
```

### **Popularity**

Popularity was based on user interactions on the TMDb website. Things like page views, users adding it to a list, users rating it, etc. Now take an overview on the popularity:

```{r}
summary(df$popularity)
```

```{r, fig.width=8}
# visualize total popularity by release year
grid.arrange(
  df %>% group_by(release_year) %>%
  summarize(total_popularity = sum(popularity)) %>%
  ggplot(aes(release_year, total_popularity)) +
  geom_point(color = "steel blue") +
    labs(title = "total popularity accross time",
         x = "release year", y = "total popularity"),

df %>% group_by(release_year) %>%
  summarize(average_p = mean(popularity),
            max_p = max(popularity),
            min_p = min(popularity)) %>%
  ggplot(aes(release_year, average_p)) +
  geom_point(color = "steel blue", alpha = 1) +
  geom_errorbar(aes(ymax = max_p, ymin = min_p, color = release_year)) +
  theme(legend.position = "none")+
  labs(title = "popularity range across time",
       x = "release year", y = "popularity"),
ncol = 2)
```

The figure total popularity across time showed that the total popularity increased by the time and its trend was similar to the trend of revenue and/or budget vs release year. 
The range of popularity also seem to be wider by the time.

**Feature engineering**

To reduce the change across time, I added one more feature `normalized_popularity` by dividing `popularity` by the average popularity per release year. The calculation is as following:

```{r}
# create the summarized table for popularity
popularity_sum <- df %>%
  group_by(release_year) %>%
  summarize(avg_popularity = mean(popularity, na.rm = T))
# create new normallized popularity
df <- df %>%
  left_join(popularity_sum, by = "release_year") %>%
  mutate(normallized_popularity = popularity/avg_popularity)
```

Adding second time normallized popularity feature.

```{r}
# create 2nd summarize table for normallized_popularity
norm_popularity_sum <- df %>%
  group_by(release_year) %>%
  summarize(max_norm_pop = max(normallized_popularity, na.rm = T))
# add 2nd normallized popularity
df <- df %>%
  left_join(norm_popularity_sum, by = "release_year") %>%
  mutate(secon_norm_popularity = normallized_popularity/max_norm_pop)
```

### **ROI**

Return on investment is a ratio between net profit and cost of investment. A high ROI means the investment's gains compare favorably to its cost. As a performance measure, ROI is used to evaluate the efficiency of an investment or to compare the efficiencies of several different investments.

```{r}
df %>% mutate(ROI = (revenue - budget)/budget) %>%
  select(title, budget, normallized_popularity, revenue, ROI, vote_count, vote_average, popularity) %>%
  arrange(desc(ROI)) %>% head(10) %>%
  kable(caption = "Top movies with greatest ROI", digits = 2) %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"), 
                position = "center", font_size = 10,
                bootstrap_options = c("striped", "hover"))
```

The most profitable movie is "Paranormal Activity" (ROI 12889), which is a 2007 American supernatural horror film co-produced, written, directed, photographed and edited by Oren Peli. It centers on a young couple (Katie Featherston and Micah Sloat) who are haunted by a supernatural presence in their home. The film earned nearly $108 million at the U.S. box office and a further $85 million internationally for a worldwide total of $193 million (reference: https://en.wikipedia.org/wiki/Paranormal_Activity).

The second most profitable movie is "The Blair Witch Project". Reference from wikipedia, The Blair Witch Project grossed nearly $250 million worldwide on a modest budget of $60,000, making it one of the most successful independent films of all time (reference: https://en.wikipedia.org/wiki/The_Blair_Witch_Project).

### **Correlation between continuous variables**

Let's take a look on the correlation between `revenue` and `budget`, `popularity`, `runtime`.

```{r, warning=FALSE, fig.width=5, fig.height=5}
corrplot(cor(df[1:3000,c("revenue","budget","popularity","runtime",
                         "normallized_popularity", 
                         "secon_norm_popularity")]), type = "upper",
         method = "number",)
```

**Insight** 

* The correlation is highest between budget and revenue. It's easy to understand because when a movie production company increase the budget, they also expect the revenue increased as well. 

* The second high correlation is between normallized population and revenue. The normallized population is correlated to revenue higher than the original popularity, but the second normallized popularity is less correlated to revenue compare to other variables. 

* Runtime is also not high correlated to revenue.

## **Revenue vs discrete variables**

We have total 17 discrete variables. Some variables have no or less informative information which I decide not to analyze them in this report (`id`, `tagline`, `title`).

### **belongs_to_collection**

This variable have greatest number of NA values. There are `r 100*round(mean(is.na(df$belongs_to_collection)),4)`
% NA values and 
`r 100*round(mean(!is.na(df$belongs_to_collection)),4)`
% not-NA values. Let's take a look on the original values. 

```{r}
head(df$belongs_to_collection,5)
```

Not all strings were necessary, only the name of the collection is informative. I extracted the collection's name by following code.

```{r}
# extract collection
df$collection <- str_extract(df$belongs_to_collection, 
            pattern = "(?<=name\\'\\:\\s{1}\\').+(?=\\'\\,\\s{1}\\'poster)")
df$collection[is.na(df$collection)] <- "no collection"
# quick review on the new feature collection
head(df$collection)
```

There were total
`r length(levels(factor(df$collection)))`
collection from both train_set and test_set.
To visualize the difference between movies belong to a collection and movies not belong to any collection, I added one new feature `collection_status` to classify a movie belong to a collection or not. The following code and figure represented for this difference.

```{r, fig.width=7, fig.height=3.5}
df <- df %>% mutate(collection_status = ifelse(collection == "no collection", 0, 1))
grid.arrange(
df %>% mutate(collection_status = factor(collection_status)) %>%
  ggplot(aes(x = collection_status, fill = collection_status)) +
  geom_bar() +
  theme_classic() +
  theme(legend.position = "none")+
  scale_fill_manual(values = c("grey","steel blue")) +
  labs(title = "frequency by collection status",
       caption = "Figure - barplot & boxplot using ggplot2,
       data df"),
  
df %>% mutate(collection_status = factor(collection_status)) %>%
  ggplot(aes(x = collection_status, revenue, fill = collection_status)) +
  geom_boxplot() +
  theme_classic() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("grey","steel blue")) +
  labs(title = "boxplot of collection status",
       caption = "0 - not belong to any collection,
        1 - belongs to collection"),
ncol = 2)
```

**Insight** almost movie didn't belong to any collection, but in average, a movie belongs to a collection has higher revenue than others. 

###  **genres**

Let's take a look on the genres. 

```{r}
head(df$genres)
```

A movie can have one genre or multiple genres. Each genre represents by following text *"{'id': 35, 'name': 'Comedy'}"* and separate by comma.

There was also
`r sum(is.na(df$genres))` 
NA value. I replaced those NA values by "no genre".

```{r}
df$genres[is.na(df$genres)] <- "no genre"
```

Suppose the first genre from the genres strings was the main genre for each movie. Therefore I extracted the first genre in genres and add them to a new variable called `main_genre`.

```{r}
# create a vector with all genre levels
genres_matching_point <- "Comedy|Horror|Action|Drama|Documentary|Science Fiction|
              Crime|Fantasy|Thriller|Animation|Adventure|Mystery|War|Romance|Music|
              Family|Western|History|TV Movie|Foreign"

# extract the main genre from genres

df$main_genre <- str_extract(df$genres, genres_matching_point)
df$main_genre[is.na(df$main_genre)] <- "no genre"
```

Now figure out the difference of each main genre, by the number of movies and revenue.

```{r, fig.width=8, fig.height=3.5}
grid.arrange(
  
# barplot number of movies per genre  
df %>% 
  group_by(main_genre) %>%
  summarize(count = n()) %>%
  mutate(main_genre = reorder(main_genre, count),
         top_genre = ifelse(main_genre %in% c("Drama","Comedy","Action"),
                            "top count",
                            ifelse(main_genre %in% c("Adventure","Animation",
                            "Science Fiction"),"top revenue","common"))) %>%
  ggplot(aes(main_genre, count, fill = top_genre)) +
  geom_col() +
  scale_fill_manual(values = c("grey","steel blue","#FF9900"))+
  theme_classic() +
  coord_flip()+
  labs(title = "counting number of first genre") +
  theme(legend.position = "none"),

# barplot average revenue per genre
df %>% 
  group_by(main_genre) %>%
  summarize(avg_revenue = mean(revenue, na.rm = TRUE)) %>%
  mutate(main_genre = reorder(main_genre, avg_revenue),
         top_genre = ifelse(main_genre %in% c("Drama","Comedy","Action"),
                            "top count",
                            ifelse(main_genre %in% c("Adventure","Animation",
                            "Science Fiction"),"top revenue","common"))) %>%
  ggplot(aes(main_genre, avg_revenue, fill = top_genre)) +
  geom_col() +
  scale_fill_manual(values = c("grey","steel blue","#FF9900"))+
  theme_classic() +
  coord_flip()+
  labs(title = "average revenue of first genre") +
  theme(legend.position = "none"),
ncol = 2)
```

**Insight** Top three common genres with highest number of movies were *Drama, Comedy, Action*.

**Insight** Top three genres with greatest average revenue were *Adventure, Animation, Science Fiction*, and the average revenue of each genre are quite different compared to others.

Because a movie could have more than one genre, I created the new variable to visualize the total genres of each movie.

```{r}
df$number_genres <- str_count(df$genres, pattern = "id")
df$number_genres[is.na(df$number_genres)] <- 0
```

Now we evaluate the correlation between budget, revenue on the number of genres.

```{r, fig.width=8, fig.height=3}
grid.arrange(
  df %>% 
    ggplot(aes(number_genres)) +
    geom_bar(fill = "steel blue"),
  
  df %>% group_by(number_genres) %>%
    summarize(avg_budget = mean(budget, na.rm = TRUE)) %>%
    ggplot(aes(number_genres, avg_budget)) +
    geom_col(fill = "steel blue"),
  
    df %>% group_by(number_genres) %>%
    summarize(avg_revenue = mean(revenue, na.rm = TRUE)) %>%
    ggplot(aes(number_genres, avg_revenue)) +
    geom_col(fill = "steel blue"),
  
  ncol = 3
)
```

**Insight** In average, movie's revenue and budget inreases proportionally with the number of genres. 

### **homepage**

Let's see the top 5 rows of homepage.

```{r}
head(df$homepage, 5)
```

There are total 
`r 100*round(mean(is.na(df$homepage)),4)`
% NA values and in this variable. Do a movie with homepage have better revenue than other movies? I added one new feature `homepage_status` with 2 levels "no homepage" and "has homepage" to answer above question.

```{r}
df$homepage_status <- ifelse(is.na(df$homepage), "no homepage","has homepage")
```

```{r}
df[1:3000,] %>% ggplot(aes(homepage_status, revenue, fill = homepage_status)) +
  geom_boxplot() +
  scale_fill_manual(values = c("steel blue","grey")) +
  theme_classic()+
  theme(legend.position = "none")
```

**Insight** A movie which homepage seem to have higher revenue compare to non-homepage movies. 

### **original_language**

Create summary table of `original_language`.

```{r}
df %>% group_by(original_language) %>%
  summarize(count = n(),
            percent = 100*round(n()/nrow(df),3),
          avg_budget = mean(budget, na.rm = TRUE),
          avg_revenue = mean(revenue,na.rm = TRUE)) %>% 
  arrange(desc(count)) %>%
  head(10) %>%
  kable(caption = "summary table of top 10 original languages", digits = 0) %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"), 
                position = "center", font_size = 10,
                bootstrap_options = c("striped", "hover"))
```

There are total 
`r length(levels(factor(df$original_language)))`
levels and 
`r mean(is.na(df$original_language))`
NA values in the original_language. English (86%) is the most popular languages compare to other languages. To reduce the number of levels on this variables, I create a new ordinal feature to rank the original languages by its percent.

```{r}
df <- df %>% 
  mutate(original_language_sum = 
     ifelse(original_language %in% c("en"),5,
       ifelse(original_language %in% c("fr,hi","ru"),4,
         ifelse(original_language %in% c("es","ja","it","de","ko","zh","cn"),3,
            ifelse(original_language %in% c("af","ar","ca","is","ka","kn","mr","nb","vi"),1,2)))))
```

### **production_companies**

Let's take a look on the top 5 rows of production_companies.

```{r}
head(df$production_companies, 5) %>%
  kable(caption = "Top rows of production companies") %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"), 
                position = "center", font_size = 10,
                bootstrap_options = c("striped", "hover"))
```

There are total `r sum(is.na(df$production_companies))`
NA values, approximate
`r 100*round(mean(is.na(df$production_companies)),4)`
%. 
Each production company represent by a name (string) and an id (number). 
There were famous production companies such as Paramount Pictures, Walt Disney Pictures... and a movie could has more than one production company. 

To explore the impact of the number of production company, I firstly created a feature called `number_of_company` by the following code: 

```{r}
# calculate number of company of a movie
df$number_of_company <- str_count(df$production_companies, pattern = "\\'name\\'")

# replace NA number by median
df$number_of_company[is.na(df$number_of_company)] <- median(df$number_of_company, na.rm = TRUE)
```

Figure out the number of production company per movie and how it correlated to the revenue.

```{r,fig.height=3, fig.width=8}
# figure out the number of company per movies and revenue change
grid.arrange(
  df %>% mutate(number_of_company = factor(number_of_company)) %>%
    ggplot(aes(number_of_company)) +
    geom_bar(fill = "steel blue", color = "white") +
    labs(title = "number of production company"),
  
df %>% 
  ggplot(aes(number_of_company, revenue)) +
  geom_point(color = "steel blue", alpha = 0.7) +
  labs(title = "revenue by number of production company"),

ncol = 2)
```

**Insight**
`r 100*round(mean(df$number_of_company <=5),3)`
% movies have less than or equal 5 production companies. The revenue seem to decrease when the number of production company greater than 5.

To explore the difference by production companies, I firstly created a feature named `companies`, contain all production company of a movie, created by following code:

```{r}
# remove all unnecessary character and keep only production company's name
df$companies <- gsub("(\\[?\\{\\'name\\'\\:\\s\\')|(\\'\\,\\s{1}\\'id\\'\\:\\s{1}\\d+\\}\\]?)","",df$production_companies)

# replace NA value in feature companies by "no production companies info"
df$companies[is.na(df$companies)] <- "no production companies info"

# create a list of all production companies
production_companies <- strsplit(df$companies, ", ") 
production_companies <- unlist(production_companies, use.names=FALSE)
```

Suppose the first company in production company list is the main company of a movie, I extracted its name and assigned to a new feature named first_company.

```{r}
# extract first company
df$first_company <- gsub("\\,\\s{1}.*","",df$companies)
```

To evaluate performance of a production company, I used following metric: 

* *number of release movies*, represent for the experience of a production company, to order a production company by its size.

* *percent*  by number of release movies, represent for the size of production company in the movie market, to order a production company by its size.

* *average budget per company*, represent for the financial power of production company.

* *average revenue per company*, represent for its average revenue.

* *revenue percent*, represent for the percent of total revenue of production company compare to other companies, to order a production company by its successful levels.

* *ROI per company*, represent for the average Return on Investment of a production company, to order a production company by its effectiveness.

```{r}
# calculate the total revenue from train data
total_revenue <- sum(train_set$revenue)

# create a summary table by first company
first_company_summary <- df[] %>% group_by(first_company) %>%
  summarize(movies_per_company = n(),
            percent = 100*round(n()/nrow(df),3),
            avg_budget_per_company = mean(budget, na.rm = TRUE),
            avg_revenue_per_company = mean(revenue, na.rm = TRUE),
            revenue_percent = round(100*sum(revenue, na.rm = TRUE)/total_revenue,0),
            ROI_per_company = round((mean(revenue, na.rm = TRUE) - 
                           mean(budget, na.rm = TRUE))/
                     mean(budget, na.rm = TRUE),3))
```

```{r}
# visualize summary table of first production company by number of release movies
first_company_summary %>% 
  arrange(desc(movies_per_company)) %>% head(10) %>%
  kable(caption = "Top production companies by number of release movies") %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"), 
                position = "center", font_size = 10,
                bootstrap_options = c("striped", "hover"))
```

**Insight** 

Although there were total 3000 movies in train data set and 
`r length(levels(factor(df$first_company[1:3000])))`
production companies in variables first company, but:

* Top 10 companies appeared as the first company in 
`r 167 + 158 + 122 + 90 + 70 + 69 + 62 + 44 + 44`
movies, approximate
`r round((167 + 158 + 122 + 90 + 70 + 69 + 62 + 44 + 44)/3000,3)*100`
%.

* only "Universal Pictures", "Paramount Pictures" and "Walt Disney Pictures" contribute approximate 27% of total revenue.

* 2.1% movies which were missing Production company information, had negative ROI (-0.547).

```{r}
top_production_companies <- first_company_summary %>% 
  arrange(desc(movies_per_company)) %>% head(10) %>%
  pull(first_company)
top_production_companies
```



### **production_countries**

Each country represent by following text "{'iso_3166_1': 'US', 'name': 'United States of America'}" and separate by comma.

```{r}
head(df$production_countries, 10)
```

Now we remove all unnecessary strings and keep only the country's name in the new feature `country`.

```{r}
# extract the first production country
df$country <- gsub("(\\[?\\{\\'iso\\_3166\\_1\\'\\:\\s{1}\\')|(\\'\\,\\s{1}\\'name.*\\}\\]?)","",df$production_countries)
df$country[is.na(df$country)] <- "no country info"

# create summary table by first production country
df %>% mutate(country = factor(country)) %>%
  group_by(country) %>%
  summarize(count = n()) %>% arrange(desc(count)) %>%
  head(15) %>%
  ggplot(aes(x = reorder(country, count), count, fill = count)) +
  labs(x = "country", title = "movies by production country") +
  geom_col() + coord_flip() + theme(legend.position = "none")
```

`r round(100*mean(df$country == "US"),1)`
movies produced in United State of America.

### **spoken_languages**

Let's take a look on top rows of `spoken_languages`.

```{r}
df$spoken_languages %>% head(10)
```

Each spoken language represent by following text "{'iso_639_1': 'en', 'name': 'English'}". We firstly create new feature `n_spoken_languages` represent for the number of spoken languages of a movie. Then we replace all unnecessary strings in the `spoken_languages`.

```{r}
# calculate number of spoken languages per movies
df$n_spoken_languages <- str_count(df$spoken_languages, pattern = "name")

# replace NA value by median
df$n_spoken_languages[is.na(df$n_spoken_languages)] <- median(df$n_spoken_languages, na.rm = T)
```

```{r, fig.width=8}
grid.arrange(
  df %>% mutate(n_spoken_languages = factor(n_spoken_languages)) %>%
  ggplot(aes(n_spoken_languages)) +
  geom_bar(fill = "steel blue", color = "white") +
    theme_classic(),
  
  df %>% mutate(n_spoken_languages = factor(n_spoken_languages)) %>%
  ggplot(aes(n_spoken_languages, budget)) +
  geom_boxplot(fill = "light blue") +
    theme_classic(),
df %>% mutate(n_spoken_languages = factor(n_spoken_languages)) %>%
  ggplot(aes(n_spoken_languages, revenue)) +
  geom_boxplot(fill = "light blue")+
  theme_classic(),
ncol = 3)
```

`r 100*round(mean(df$n_spoken_languages == 1),3)`
% movies have one spoken languages, another
`r 100*round(mean(df$n_spoken_languages == 2),3)`
% movies have two spoken languages, only 
`r 100 - 100*round(mean(df$n_spoken_languages <=2),3)`
% movies have more than two spoken languages.

### **cast**

A movie could have many casts, each `cast` represented by following text "{'cast_id': 4, 'character': 'Lou', 'credit_id': '52fe4ee7c3a36847f82afae7', 'gender': 2, 'id': 52997, 'name': 'Rob Corddry', 'order': 0, 'profile_path': '/k2zJL0V1nEZuFT08xUdOd3ucfXz.jpg'}".

To explore the correlation between number of cast per movies and the revenue, I added a new feature named n_cast to summarize the number of cast of a movie. In case missing value, I replaced by median.

```{r}
# create feature n_cast to summarize number of cast per movie
df$n_cast <- str_count(df$cast, pattern = "name")

# replace NA value by median.
df$n_cast[is.na(df$n_cast)] <- median(df$n_cast, na.rm = T)
```

```{r, fig.height=3.5, fig.width=3.5}
corrplot(cor(select(df[1:3000,], revenue, budget, n_cast)),
         type = "upper", method = "number")
```

As we saw, the correlation between number of cast vs budget and/or revenue was not really high (0.3 and 0.34).

**Specific effect by leading actor/actress**

Normally in each movie, there was a leading actor, leading actress, star, or simply lead who plays the role of the protagonist of a film, television show or play. The word lead may also refer to the largest role in the piece and leading actor may refer to a person who typically plays such parts or an actor with a respected body of work.

We also know in practice, many people watch a movie just because the appearance of their idols or favor actors. To explore the correlation between the "leading actor/actress", I extracted name of the first and second casts into two feature `first_cast` and `second_cast`. 

```{r}
# extract the first cast
df$first_cast <- gsub("\\'\\,\\s\\'order\\'\\:\\s+0\\,.*",
                      "",
                      df$cast)
index <- str_which(df$first_cast,"profile")
df$first_cast <- gsub("\\'\\,\\s\\'order\\'\\:\\s+1\\,.*",
                      "",
                      df$first_cast)
df$first_cast <- str_extract(df$first_cast,"(?<=name\\'\\:\\s\\').+")
```

```{r}
# extract the second cast
df$second_cast <- gsub("\\'\\,\\s\\'order\\'\\:\\s{1}1\\,.*",
                      "",
                      df$cast)
df$second_cast <- gsub("\\[.*name\\'\\:\\s\\'","",
                       df$second_cast)
df$second_cast[index] <- gsub("\\'\\,\\s\\'order\\'\\:\\s{1}2\\,.*",
                      "",
                      df$cast)
df$second_cast[index] <- gsub("\\[.*name\\'\\:\\s\\'","",
                       df$second_cast)
```

```{r}
# create a summary table of first and second cast
cast_summary <- df %>% 
  select(first_cast, second_cast, revenue, budget) %>%
  gather(1:2, key = "cast_order", value = "cast_name")

cast_summary$cast_name <- gsub("\\'\\,.*","",cast_summary$cast_name)

cast_summary <- cast_summary %>%
  group_by(cast_name) %>%
  summarize(movies_per_cast = n(),
            avg_revenue_per_cast = mean(revenue, na.rm = T))

# remove all NA values
cast_summary <- na.omit(cast_summary)
```

```{r}
# visualize table
cast_summary %>% arrange(desc(movies_per_cast)) %>%
  head(10) %>%
  kable(caption = "Top 10 leading actor/actress all the time") %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"), 
                position = "center", font_size = 10,
                bootstrap_options = c("striped", "hover"))
```

Top 3 actor/actress with greatest number of movies were Robert De Niro, Nicolas Cage, Bruce Willis. The median revenue per cast seem to be increased when the number of movies increase.

```{r, fig.width=7, fig.height=5}
grid.arrange(
cast_summary %>% mutate(movies_per_cast = factor(movies_per_cast)) %>%
  ggplot(aes(movies_per_cast)) +
  geom_bar(fill = "steel blue", color = "white") +
  labs(title = "frequency by number of movies per cast"),
  
cast_summary %>% mutate(movies_per_cast = factor(movies_per_cast)) %>%
  ggplot(aes(movies_per_cast, avg_revenue_per_cast)) +
  geom_boxplot(fill = "light blue") + 
  labs(title = "average revenue vs number of movies per cast"),

ncol = 1)
```

There were
`r 100*round(mean(cast_summary$movies_per_cast <=5),3)`
% cast appeared less than or equal 5 times, 
`r 100*round(mean(cast_summary$movies_per_cast <=10) - mean(cast_summary$movies_per_cast <=5),3)`
% cast appeared more than 5 and less than or equal 10 times, only
`r 100 - 100*round(mean(cast_summary$movies_per_cast <=10),3)`
% cast appeared more than 10 times.
Although the median seem to be increased when the number of movies per cast increase, the range of average revenue is so wide

### **crew**

Crew contain the production team of a movie, each member represents by following strings "{'credit_id': '59ac067c92514107af02c8c8', 'department': 'Directing', 'gender': 0, 'id': 1449071, 'job': 'First Assistant Director', 'name': 'Kelly Cantley', 'profile_path': None}" and separate with other members by comma.

Because the length of strings is too long, I didn't copy it to this report.

Let's see how many members of a movies and explore its correlation to the movie's revenue.

```{r}
# calculate the number of crew per movie
df$n_crew <- str_count(df$crew, pattern = "name")

# replace NA value by median
df$n_crew[is.na(df$n_crew)] <- median(df$n_crew, na.rm = T)
```

```{r, fig.width=7, fig.height=3}
# visualize the frequency of number of crew and correlation between n_crew and revenue
grid.arrange(
  df %>% ggplot(aes(n_crew)) +
  geom_histogram(fill = "steel blue", color = "white") +
    labs(title = "number of crew"),
  
  df %>% ggplot(aes(n_crew, revenue)) +
    geom_point(color = "steel blue") +
    geom_smooth() + 
    labs(title = "revenue vs number of crew"),
  
  ncol = 2)
```

Looked like that there were low correlation between number of crew and revenue. The correlation was 
`r cor(df$revenue[1:3000], df$n_crew[1:3000])` which is not a high correlation.


**Effect by director**

A film director controls a film's artistic and dramatic aspects and visualizes the screenplay (or script) while guiding the technical crew and actors in the fulfilment of that vision. The director has a key role in choosing the cast members, production design, and the creative aspects of filmmaking.(*Reference: https://en.wikipedia.org/wiki/Film_director*)

To evaluate impact from a director, we extract the director's name from `crew` by following code:

```{r}
# extract the director name
df$director <- gsub(".*\\'Director\\'\\,\\s\\'name\\'\\:\\s\\'","",df$crew)
df$director <- gsub("\\'\\,\\s.*","",df$director)
```

Create a summary table for director.

```{r}
# create summary table by director
director_summary <- df %>% select(director, revenue) %>%
  group_by(director) %>%
  summarize(movies_per_director = n(),
            avg_revenue_per_director = mean(revenue, na.rm = T))
```

There were total
`r nrow(director_summary)`
directors in our data. To explore the correlation between director experience vs revenue.

```{r, fig.width=8, fig.height=3}
grid.arrange(
  director_summary %>% mutate(movies_per_director = factor(movies_per_director)) %>%
  ggplot(aes(movies_per_director)) +
  geom_bar(fill = "steel blue", color = "white"),
  
  director_summary %>% mutate(movies_per_director = factor(movies_per_director)) %>%
  ggplot(aes(movies_per_director, avg_revenue_per_director)) +
  geom_boxplot(),
  
  ncol = 2)
```


```{r}
# create top 10 directors with highest number of movies
director_summary %>% arrange(desc(movies_per_director)) %>%
  head(10) %>%
  kable(caption = "Top 10 directors with greatest number of movies") %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"), 
                position = "center", font_size = 10, full_width = F,
                bootstrap_options = c("striped", "hover"))
```

Top 10 directors with greatest revenue.

```{r}
director_summary %>% arrange(desc(avg_revenue_per_director)) %>%
  head(10) %>% 
  kable(caption = "Top 10 directors with greatest revenue per movies") %>%
  kable_styling(latex_options = c("HOLD_position","scale_down"), 
                position = "center", font_size = 10, full_width = F,
                bootstrap_options = c("striped", "hover"))
```

## **Features engineering**

### **collection revenue**

From previous section, we know movies belong to a collection have greater revenue than moves not belong to any collection. We will calculate the first new feature as average revenue per each collection.

```{r}
collection_revenue <- df %>% group_by(collection) %>%
  summarize(avg_revenue_per_collection = mean(revenue, na.rm = T))
```

The feature `aveg_revenue_per_collection` have 
`r 100*round(mean(is.na(collection_revenue$avg_revenue_per_collection)),4)`
% NA values in total. We replace it by the median value.

```{r}
index <- is.na(collection_revenue$avg_revenue_per_collection)
collection_revenue$avg_revenue_per_collection[index] <- 
  median(collection_revenue$avg_revenue_per_collection, na.rm = T)
```

```{r}
df <- df %>% left_join(collection_revenue, by = "collection")
```

### **Production companies, cast, director**

Now we add the effect of production companies, casts and directors to our model.

**Production company**

```{r}
# left join the summary features by company
df <- df %>% left_join(first_company_summary[,c(1,2,5,6)], by = "first_company")
```

**First cast and second cast**

```{r}
# change cast_name to first_cast
names(cast_summary) <- c("first_cast","movies_per_cast","avg_revenue_per_cast")
# left join by first_cast
df <- df %>% left_join(cast_summary[,c(1,3)], by = "first_cast")
# change cast_name to second_cast
names(cast_summary) <- c("second_cast","movies_per_cast","avg_revenue_per_cast")
# left join by second cast
df <- df %>% left_join(cast_summary[,c(1,3)], by = "second_cast")
```

**Director**

```{r}
# left join by director
df <- df %>% left_join(director_summary, by = "director")
```

**Genres**

```{r}
genres <- levels(factor(df$main_genre))
```

```{r}
for (i in 1:19){
  df[,genres[i]] <- ifelse(str_detect(df$genres,genres[i]),
                           1,0)
}
```

**Top production companies**

```{r}
grep("no production companies info", colnames(df))
```


```{r}
for (i in 1:10){
  df[,top_production_companies[i]] <- 
    ifelse(str_detect(df$production_companies,top_production_companies[i]),
                           1,0)
}

df$`no production companies info` <-
  ifelse(str_detect(df$companies, "no production companies info"),1,0)

from <- grep("no production companies info", colnames(df))

to <- grep("Columbia Pictures Corporation", colnames(df))

df$other_production_company <-
  ifelse(rowMeans((df[,from:to])) >0,0,1)

for (i in 87:(ncol(df))){
  df[is.na(df[,i]),i] <- 0
}
```

Selecting the features.

```{r selecting features,}
dat <- df %>% 
    mutate(x2 = budget*normallized_popularity,
         x3 = budget*vote_average,
         x4 = budget*secon_norm_popularity) %>%
  select(revenue,
         x2,budget,
         #avg_revenue_per_director,
         #avg_revenue_per_collection,
         #avg_revenue_per_cast.x,
         #avg_revenue_per_cast.y,
         normallized_popularity, 
         collection_status,
         release_year, release_month, 
         popularity,
         number_of_company,
         n_spoken_languages, original_language_sum,
         #n_crew, n_cast,
         number_genres,
         68:79,81:83,85:94,96:97)
```


```{r}
head(dat)
```


Review the NA values.

```{r}
plot_missing(dat)
```

Replace NA values by median.

```{r}
for (i in 2:(ncol(dat))){
  dat[is.na(dat[,i]),i] <- median(dat[,i], na.rm = T)
}
```

**Correlation plot**

```{r, fig.width=7, fig.height=7}
corrplot(cor(dat[1:3000,]), type = "upper", method = "ellipse")
```

# **Modeling**

Before process modeling, we transform the revenue, budget, av_revenue_per_director, x2, x4 by logarit 10.


```{r}
for (i in 1:3){
  dat[,i] <- logb(dat[,i]+1)
}

head(dat)
```

```{r}
dat$avg_revenue_per_director <- dat$avg_revenue_per_director/mean(dat$avg_revenue_per_director)
dat$avg_revenue_per_collection <- dat$avg_revenue_per_collection/mean(dat$avg_revenue_per_collection)
dat$avg_revenue_per_cast.x <- dat$avg_revenue_per_cast.x/mean(dat$avg_revenue_per_cast.x)
dat$avg_revenue_per_cast.y <- dat$avg_revenue_per_cast.y/mean(dat$avg_revenue_per_cast.y)
```


We use randorm forest (rf), support vector machine (svmLinear), gradient linear model boosting (glmboost) and linear model (lm) for modeling.

```{r}
fit_rf <- train(revenue ~ ., 
             data = dat[1:3000,],
             method = "rf",
             metric = "RMSE",
             trControl = trainControl(method = "cv",
                                      number = 5,
                                      p = 0.8))
fit_svm <- train(revenue~.,
                 data = dat[1:3000,],
                 method = "svmLinear",
                 metric = "RMSE",
                 trControl = trainControl(method = "cv",
                                      number = 5,
                                      p = 0.8))
fit_glmboost <- train(revenue~.,
                 data = dat[1:3000,],
                 method = "brnn",
                 metric = "RMSE",
                 trControl = trainControl(method = "cv",
                                      number = 5,
                                      p = 0.8))
fit_lm <- train(revenue~.,
                 data = dat[1:3000,],
                 method = "bayesglm",
                metric = "RMSE",
                 trControl = trainControl(method = "cv",
                                      number = 5,
                                      p = 0.8))
```

```{r}
data.frame(method = c("rf","svm","glmboost","lm"),
           RMSLE =c(min(fit_rf$results$RMSE),
                    min(fit_svm$results$RMSE),
                    min(fit_glmboost$results$RMSE),
                    min(fit_lm$results$RMSE)),
           Rsquared = c(max(fit_rf$results$Rsquared),
                        max(fit_svm$results$Rsquared),
                        max(fit_svm$results$Rsquared),
                        max(fit_lm$results$Rsquared)))

```

```{r}
# calculate predicted revenue
yhat_rf <- predict(fit_rf, newdata = dat[3001:7398,])
yhat_svm <- predict(fit_svm, newdata = dat[3001:7398,])
yhat_glmboost <- predict(fit_glmboost, newdata = dat[3001:7398,])
yhat_lm <- predict(fit_lm, newdata = dat[3001:7398,])
```

```{r}
# ensembles
ensembles <- data.frame(rf = yhat_rf,
                        svm = yhat_svm,
                        glmboost = yhat_glmboost,
                        lm = yhat_lm)
```

```{r}
corrplot(cor(ensembles), method = "number", type = "upper",na.label.col = TRUE)
```

```{r}
ensembles <- ensembles %>% mutate(combine = rowMeans(ensembles)) %>%
  mutate(combine = ifelse(combine <= 0|combine == Inf, 7, combine))
```

```{r}
RMSE <- function(predicted_revenue, true_revenue){
  sqrt(mean((true_revenue- predicted_revenue)^2))
}
```


```{r}
results <- data.frame(rf = RMSE(ensembles$rf, logb(test_y)),
           svm = RMSE(ensembles$svm, logb(test_y)),
           glmboost = RMSE(ensembles$glmboost, logb(test_y)),
           lm = RMSE(ensembles$lm, logb(test_y)),
           ensemble = RMSE(ensembles$combine, logb(test_y)))
results           
```

```{r}
ensembles <- ensembles %>%
  mutate(true_revenue = logb(test_y))

cor(ensembles)
```

```{r}
ensembles %>% mutate(error = true_revenue - combine) %>%
  ggplot(aes(true_revenue, error)) +
  geom_point(color = "steel blue", alpha = 0.3)
```


```{r}
ensembles %>% mutate(true_revenue = logb(test_y)) %>%
  ggplot() +
  geom_density(aes(true_revenue), fill = "gray", alpha = 0.3) +
  geom_density(aes(rf), fill = "blue", alpha = 0.3) +
  geom_density(aes(svm), fill = "green", alpha = 0.3) +
  geom_density(aes(lm), fill = "orange", alpha = 0.3)
  
```

# **Modeling subset by year**

```{r}
release_years <- levels(factor(df$release_year))
```

```{r}
dat_test <- dat[3001:7398,] %>%
  mutate(predicted_revenue = ensembles$combine,
         true_revenue = test_y) %>%
  mutate(error = test_y - predicted_revenue)
dat_test %>% ggplot(aes(release_year, error)) +
  geom_point(color = "steel blue")
```

```{r}
years_group <- c(1,2,3)

control <- trainControl(method = "cv",
                        number = 5,
                        p = 0.75)
dat <- dat %>% 
  mutate(yr_group = ifelse(release_year <= 1990,1,
                           ifelse(release_year > 1990 & release_year <= 1990,2,
                                  ifelse(release_year > 1990 & release_year <= 2010, 2,3))))

fit_60s_90s <- train(revenue~.,
                     data = filter(dat[1:3000,], yr_group == 1),
                     method = "rf",
                     metric = "RMSE",
                     trControl = control)

fit_90s_2k10 <- train(revenue~.,
                     data = filter(dat[1:3000,], yr_group == 2),
                     method = "rf",
                     metric = "RMSE",
                     trControl = control)
fit_after_2k10 <- train(revenue~.,
                     data = filter(dat[1:3000,], yr_group == 3),
                     method = "rf",
                     metric = "RMSE",
                     trControl = control)
yhat <- dat[3001:7398,] %>% 
  mutate(yhat_2 = predict(fit_60s_90s, dat[3001:7398,]),
         yhat_3 = predict(fit_90s_2k10, dat[3001:7398,]),
         yhat_4 = predict(fit_after_2k10,dat[3001:7398,])) %>%
  mutate(final_revenue = ifelse(yr_group == 1, yhat_2,
                                ifelse(yr_group == 2, yhat_3, yhat_4))) %>%
  pull(final_revenue)
RMSE(yhat, logb(test_y))
```

```{r}
cor(yhat, logb(test_y))
```


```{r}
dat[3001:7398,] %>%
  mutate(predicted_revenue = yhat,
         true_revenue = test_y) %>%
  mutate(error = true_revenue - predicted_revenue)%>% 
  ggplot(aes(release_year, error)) +
  geom_point(color = "steel blue", alpha = 0.3)
```


```{r}
data.frame(predict = yhat, true_revenue = log10(test_y)) %>%
  ggplot(aes(true_revenue, predict)) +
  geom_point(color = "steel blue", alpha = 0.5) +
  geom_smooth()
```

## **Validation**

## **Discussion**

# **Conclusion**

```{r}
head(addition_data)
```


# **Reference**